### 2024-01-31 ms.qmd
Updates

```{r}
#| label: setup
pacman::p_load(tidyverse, flextable, emmeans, DHARMa, brms, here, ggplot2, lme4, zoo, lmerTest, broom, tidybayes)
```

## Introduction

Cognition is defined as the processes by which animals gather, preserve, and use information from their environment through perception, learning, memory, and decision making [@shettleworth]. These cognitive processes underpin several aspects of animals' ecology such as foraging, mate choice, antipredatory strategies, and/or social behaviours, that are crucial for the survival and reproduction of animals [@dukas_evolutionary_2004]. Particularly, learning - the acquisition of neuronal representations of new information [@dukas_evolutionary_2004] - is seen as fundamental for coping with environmental changes by enabling individuals to create new associations between events [@dukas_evolutionary_2004; @leal_behavioural_2012; @buchanan_condition_2013]. However, in nature, there is individual variation in the capacity to acquire new information, influenced by factors like age, sex, gut microbiota, or the environment where they develop [@szuran_water_1994; @lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @carazo_sex_2014; @noble_age-dependent_2014; @alemohammad_2022_microbiota_learning]. In particular, early life conditions can play a pivotal role since the brain is especially susceptible to environmental influences during early stages of development [@zhu_prenatal_2004]. Therefore, investigating the effects of the developmental environment on learning can be essential to understand how individual variation in learning is shaped, and, in consequence, to predict animals' responses towards environmental change.  
In this sense, prenatal Glucocorticoids (GCs) - hormones related to organisms' response to stress [@sapolsky_how_2000] - and prenatal thermal environment are known to influence learning abilities in different taxa [see @lemaire_prenatal_2000; @zhu_prenatal_2004; @amiel_hotter_2012; @amiel_egg_2014; @crino_corticosterone_2014; @abayarathna_effects_2020]. For instance, some studies have demonstrated that prenatal stress and high prenatal GC levels impair learning [@lemaire_prenatal_2000; @zhu_prenatal_2004; @farrell_developmental_2015-learn], while others showed diverse effects depending on factors like subjects' sex or the nature of the learning task [@szuran_water_1994; @crino_corticosterone_2014-learn; @farrell_developmental_2015-learn; @bebus_associative_2016]. Similarly, some experiments have shown significant impacts of prenatal temperature on learning in ectotherms [@amiel_hotter_2012; @amiel_egg_2014; @dayananda_incubation_2017; @abayarathna_effects_2020]. For example, high incubation temperatures have been linked with faster learning rates in skinks [@amiel_hotter_2012; @amiel_egg_2014], while velvet geckos incubated at temperatures over their natural range learn slower than those incubated within the natural thermal limits [@abayarathna_effects_2020]. In this vein, the effect of prenatal temperature appears to be linked to alterations in neural structure and metabolic activity [@coomber_independent_1997; @sakata_neural_2000; @amiel_effects_2017; @beltran_are_2021] that share some similarities with those resulting from prenatal increased stress or GC levels [@lemaire_prenatal_2000; @zhu_prenatal_2004; @du_dynamic_2009]. This suggests that prenatal GCs and temperature can act on the same physiological mechanisms and, thus, both could interact to shape individual variation in learning abilities [@noble_developmental_2018]. Furthermore, GCs can play an pivotal role in determining vertebrate responses to elevated temperatures [@Crino_2023] potentially fostering natural interactions between temperature and GCs. However, despite the proximate similarities of prenatal GCs and temperature effects, and the potential role of GCs in vertebrates' response to elevated temperatures, our understanding of how these two factors interact remains incomplete.  
In this study, our objective is to explore the interactive effects between prenatal Glucocorticoids (GCs) and the prenatal thermal environment on learning. We utilized two species of skinks, the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_), as model species. We experimentally increased Corticosterone (CORT) - the main GC in birds, reptiles, amphibians, and rodents [@Crino_2023] - levels in the eggs of these two species of skinks and then incubated them at two different temperatures in a 2X2 factorial design. Post-incubation, the juveniles were subjectd to a colour-associative and a reversal task to assess their learning abilities. Our hypothesis posits that changes in CORT levels and temperature during early development will induce sustained effects on brain's physiology that will ultimately impact learning skills. We predict that individuals exposed to high levels of CORT and/or low temperatures will perform less proficiently in both learning tasks compared to control individuals or those exposed to high temperatures. Additionally, we anticipate that incubation at high temperatures will mitigate the impact of CORT on skink performance, while cold incubation temperatures are expected to enhance the detrimental effects of CORT on learning. We also expect the treatments to affect both tasks equally, with those individuals exposed to high levels of CORT and/or low temperatures performing less proficiently in both tasks compared to control individuals or those exposed to high temperatures. Finally, we expect that the effects of the treatments will be similar in both species, as both species share similar life history traits and are closely related [@chapple_know_2011; @chapple_biology_2014], and other cognitive studies have not found any difference between species when tested in an associative learning task [@bezzina2014does].

@@

We employed juveniles of both species, _L. delicata_ and _L. guichenoti_ that came from a breeding colony kept in the lab (details of breeding colony and husbandry in Supplementary Material). Those individuals were treated with prenatal CORT or vehicle, and incubated at different temperatures (see Supplementary Material), but previous data demonstrated that there was no differences in learning skills caused by these treatments (**REF**).
To estimate behavioural flexibility, we examined lizards' performance in a learning task and its reversal (**REF**) using colour as the main conditioned stimulus and food as a reward (see @fig-Methods). There was first a habituation phase where lizards had to learn to eat from white 3D-printed PLA ramps (9 L x 4 W x 5 H cm); then a second phase where lizards had to associate the attainability of food with one colour; and finally a reversal task where the colour associated with attainable food was changed (@fig-Methods B). We evaluated the correlation between performance in the associative and reversal task as the main estimate of behavioural flexibility.  
The habituation phase was divided into three stages: in the first stage, the reward, a small, frozen cricket (*A. domestica*) was placed in an opaque petri dish (3D x 1.6H cm) (@fig-Methods B, Stage 1); in the second stage, the petri dish with the cricket was placed on top of the white 3D printed ramps (@fig-Methods B, Stage 2); and in the third stage, the cricket was left inside a well (3D x 1.75H cm) on the top of the ramp (Fig. @fig-Methods B, Stage 3). Trials in every stage lasted one hour, begining when the feeding block (petri dish, ramp, or both) was placed inside the enclosure and finished when we took the feeding block away. We recorded whether the cricket had been consumed or not and we considered an stage as completed if the lizards ate the crickets in at least 5 out of 6 trials. This phase lasted 38 days, and only in one case we decided not to use the lizard in the next phases because its behaviour was not consistent.  
In the associative phase, we used three ramps that were identical in shape and size but differ in the colour (Associative task in Fig. @fig-Methods B). As in the last stage of the habituation phase, we placed the cricket inside the wells of the ramps, but then two of them were covered with 3D-printed lids (3D x 0.5H cm) so prey was only accessible in "the correct" ramp. The food reward was placed in all three wells to avoid lizard using prey chemical cues, that were released in the closed wells thanks to a series of small holes on the top of the lids. The colours of the feeders were green, red, and blue, as previous studies demonstrate that squamates can discriminate between these colours [@Baden_Osorio_2019_Vert_vision]. However, we decided to split the lizards in two groups to control for potential colour preference that could bias our results. One group of lizards was assigned the blue ramp as the correct choice (i.e., the ramp with the non-covered frozen cricket) in the associative task and the red ramp in the reversal, while the other group was assigned red as correct for the associative and blue for the reversal. In all trials, the position of the feeders was changed randomly to ensure subjects were using colour rather than spatial cues for the association. Lizards were tested in this task once a day for 35 days. After the colour association phase, we performed a choice reversal task (Reversal task in @fig-Methods B). Here, the attainability of prey was indicated by a different colour from the reversal, requiring the lizards to form a novel association between the new colour and the food reward. This test was done once a day for 40 days.  
We performed the full experiment was between the 6^th^ of March until the 26^th^ of June 2023, between 11 to 12 am, when the lizards were active. Trials in the learning phases (colour associative task and reversal tasks) were recorded with different CCTV systems always using the same camera per individual. Videos were analysed manually using a standard video player (IINA) by PR, who recorded whether the first choice made by each subject was the correct feeder or not. A choice was considered to be made if the head of the lizard was inside the well of one of the ramps. We considered a trial failed if there was no choice in one hour of recording and those trials were considered as 'non data' in the analyses. We excluded from our analyses those individuals with more than 15 trials failed (i.e. they did not make a choice), and we considered the first trial to be the first one where the individual made a choice. We also considered individuals to be reinforced if the cricket was eaten after the test independently of whether their initial choice was correct or not.

@@

knitr::include_graphics("./Others/LEARN_FIG_1.svg")

@@

**As such, we run a total of different 8 models employing Bayesian multilevel models using the brm function from the brms package (@burkner2017brms) in Quarto (@Allaire_Quarto_2022). Each model consisted of four parallel chains of 3000 iterations, with a warm up interval of 1000 iterations.'Choice', i.e. whether the individual chose correct (1) or not (0) was used as the response variable. The fixed effects of the model included a triple interaction between: trial ('Associative trial' or 'Reversal trial') as a numeric variable, and hormone treatmet ('CORT' versus 'Control') and the temperature at which eggs were incubated ('Cold' versus 'Hot') as factors. For the random effects, we employed lizard identity as a random intercept, and as a random slope we included the variable trial ('Associative' or 'Reversal') within each level of lizard identity. We used the resulting posterior of these models to evaluate learning differences between treatments within and between species and colour assigned. More specifically, we calculated learning slopes by using the estimates of the trial variable per each level of the hormone-temperature interaction ('Treatments'); values bigger from zero were considered as evidence of learning, while those less or equal to zero not. We used the pmcmc method to test whether those slopes or the comparissons between 'Treatments' (e.g. slope for 'CORT-Cold' lizards minus 'CORT-Hot' lizards) were different from zero (two-tailed tests). We considered statistical significance if p-value < 0.05.**

@@

>>>>

## Introduction
Behavioural flexibility, **the ability to adapt behaviour to changing environments** (**REFS**), is considered crucial for invasion success [@sol2002behavioural]. In general, species with higher behavioural flexibility are more likely to survive and establish in new habitats (**REFS**) as behaving flexibly enhances their capacity to exploit new resources and face new challenges (**REFS**).

adaptive flexibility hypothesis says that flexibility may be relevant in early and late stages of invassion (Wright et al 2010). But at the same time behavioural flexibility  it is related to relative brain size (Sol et al 2002, 2008), that may be more fixed than we thought (controversy with Wright et al??). In contrast, other hyoptheses predict this is constant, or there is individual variation in beh flexibility (Wright et al 2010). Since some aspects of early environment can play a prominent role in shaping neural substrate, we hypothesize that beh flexibility may be affected by early environment. Also, since beh flexibility is advantageous at different stages of colonization (**REFS**), invassive species may be able to maintain this behavioral flexibility through generations. In other words, there may be some mechanisms in invassive species to avoid early developmental constraints. CORT can be essential in this process as it is associated to vertebrates response to environmental changes, and it is known to impact brain development (Zhu et al 2004). Also, temperature may be relevant, since it affects brain development in ectotherms, it can be one of the main environmental changes to face by animals, andvertebrates reponse to changes in temeprature is also associated with the production of CORT (Crino et al 2023).

Here, we aim to study the effects of prenatal environment on behavioural flexibility in the invassion ecology context. To do so, we employed as our model species the delicate skink (_Lampropholis delicata_) and the common garden skink (_L. guichenoti_). These two species are closely related and share similar life history traits [@chapple_know_2011; @chapple_biology_2014], but only _L. delicata_ has been succesful in colonising different areas around the globe (**REFS**). Some studies have shown differences in personality traits between these two species that can be related to higher behavioural flexibility in _L. delicata_ (**REFS**), and are likely key for its invassive success (**REFS**). Our goal in this study was to investigate if behavioural flexibility is affected by the prenatal environment and the species specific response to it. Particularly, we were interested in the interactive effects of prenatal temperature and Corticosterone (CORT) - the main type of GC in reptiles [@Crino_2023] - because CORT is usually associated to vertebrates response to changes in temeprature [@Crino_2023], and both factors are known to impact brain development [@coomber_independent_1997; @zhu_prenatal_2004]. We hypothesized that, since behavioural flexibility is advantageous at different stages of colonization (**REFS**), invassive species may be able to maintain this behavioral flexibility through generations. In other words, there may be some mechanisms in invassive species to avoid early

@@

#### Experimental design
We employed juveniles of both species, _L. delicata_ and _L. guichenoti_ that came from a breeding colony kept in the lab (details of breeding colony and husbandry in Supplementary Material). To see the effect of prenatal CORT and temperature, eggs were topically supplied with 5µL of CORT dissolved in 70% Ethanol and 30% DMSO ('CORT' treatment) or the same volume of the vehicle ('Control'), and were then incubated at different temperatures ('Cold' -- 23ºC ± 3ºC or 'Hot' -- 30ºC ± 3ºC) until hatching.  
To estimate behavioural flexibility, we examined lizards' performance in a reversal learning task using colour as the main conditioned stimulus and food (frozen small *A. domestica*) as a reward (see @fig-Methods). The learning test was divided in three main phases: there was first a habituation phase where lizards had to learn to eat from white 3D-printed PLA ramps (9 L x 4 W x 5 H cm); then a second phase where lizards had to associate the attainability of food with one colour; and finally the reversal task where the colour associated with the attainable food was changed (@fig-Methods B). In associative and reversal tasks we used  using three different ramps that were identical in size and shape to the ones for the habituation phase. The colours of the ramps were green, red, and blue, as previous studies demonstrate that squamates can discriminate between these colours [@Baden_Osorio_2019_Vert_vision]. To control for potential colour biases, we assigned one group of lizards to associate the blue ramp with the attainable food in the associative task and the red ramp in the reversal, while the other group was assigned red as correct for the associative and blue for the reversal. In all trials, the position of the feeders was changed randomly to ensure subjects were using colour rather than spatial cues for the association. To avoid the use of chemical cues, we placed the food reward in all three ramps, but in two of them the cricket was covered by small white lids with holes on the top. Lizards were tested once daily between 11 to 12 am, when lizards are active. The training phase lasted 38 days, the Associaive learning task 35, and the Reversal learning task 40 days. We recorded and analysed manually all trials form the learning phases using a standard video player (IINA). PR was the main researcher analysing the data, and was blinded to the lizards' treatment until the end of the analyses. We used as the main variable whether the animal chose the correct ramp first or not. We considered that a choice was made if the head of the lizard was inside the well of one of the ramps. We considered a trial failed if there was no choice in one hour of recording and those trials were considered as 'non data' in the analyses. Here, we analysed the data of the reversal as a measure of behavioural flexibility. As such, we only included in our analyses those individuals who passed a learning criterion of 80% correct choices in the last 10 trials of the associative task.

@@

knitr::include_graphics("./Others/BEHFLEX_FIG_1.svg")

@@

We also considered individuals to be reinforced if the cricket was eaten after the test independently of whether their initial choice was correct or not.  
**As such, we run a total of different 8 models employing Bayesian multilevel models using the brm function from the brms package (@burkner2017brms) in Quarto (@Allaire_Quarto_2022). Each model consisted of four parallel chains of 3000 iterations, with a warm up interval of 1000 iterations.'Choice', i.e. whether the individual chose correct (1) or not (0) was used as the response variable. The fixed effects of the model included a triple interaction between: trial ('Associative trial' or 'Reversal trial') as a numeric variable, and hormone treatmet ('CORT' versus 'Control') and the temperature at which eggs were incubated ('Cold' versus 'Hot') as factors. For the random effects, we employed lizard identity as a random intercept, and as a random slope we included the variable trial ('Associative' or 'Reversal') within each level of lizard identity. We used the resulting posterior of these models to evaluate learning differences between treatments within and between species and colour assigned. More specifically, we calculated learning slopes by using the estimates of the trial variable per each level of the hormone-temperature interaction ('Treatments'); values bigger from zero were considered as evidence of learning, while those less or equal to zero not. We used the pmcmc method to test whether those slopes or the comparissons between 'Treatments' (e.g. slope for 'CORT-Cold' lizards minus 'CORT-Hot' lizards) were different from zero (two-tailed tests). We considered statistical significance if p-value < 0.05.**

@@

### Learning task
The habituation phase was divided into three stages: in the first stage, the reward, a small, frozen cricket (*A. domestica*) was placed in an opaque petri dish (3D x 1.6H cm) (@fig-Methods B, Stage 1); in the second stage, the petri dish with the cricket was placed on top of the white 3D printed ramps (@fig-Methods B, Stage 2); and in the third stage, the cricket was left inside a well (3D x 1.75H cm) on top of the ramp (Fig. @fig-Methods B, Stage 3). Trials in every stage lasted one hour, begining when the feeding block (petri dish, ramp, or both) was placed inside the enclosure and finished when we took the feeding block away. We recorded whether the cricket had been consumed or not and we considered an stage as completed if the lizards ate the crickets in at least 5 out of 6 trials. This phase lasted 38 days, and only in one case we decided not to use the lizard in the next phases because its behaviour was not consistent.  
In the associative phase, we used three ramps that were identical in shape and size but differ in the colour (Associative task in Fig. @fig-Methods B). As in the last stage of the habituation phase, we placed the cricket inside the wells of the ramps, but then two of them were covered with 3D-printed lids (3D x 0.5H cm) so prey was only accessible in "the correct" ramp. The food reward was placed in all three wells to avoid lizard using prey chemical cues, that were released in the closed wells thanks to a series of small holes on the top of the lids. The colours of the feeders were green, red, and blue, as previous studies demonstrate that squamates can discriminate between these colours [@Baden_Osorio_2019_Vert_vision]. However, we decided to split the lizards in two groups to control for potential colour preference that could bias our results. One group of lizards was assigned the blue ramp as the correct choice (i.e., the ramp with the non-covered frozen cricket) in the associative task and the red ramp in the reversal, while the other group was assigned red as correct for the associative and blue for the reversal. In all trials, the position of the feeders was changed randomly to ensure subjects were using colour rather than spatial cues for the association. Lizards were tested in this task once a day for 35 days. After the colour association phase, we performed a choice reversal task (Reversal task in @fig-Methods B). Here, the attainability of prey was indicated by a different colour from the reversal, requiring the lizards to form a novel association between the new colour and the food reward. This test was done once a day for 40 days.  
We performed the full experiment was between the 6^th^ of March until the 26^th^ of June 2023, between 11 to 12 am, when the lizards were active. Trials in the learning phases (colour associative task and reversal tasks) were recorded with different CCTV systems always using the same camera per individual. Videos were analysed manually using a standard video player (IINA) by PR, who recorded whether the first choice made by each subject was the correct feeder or not. We considered a choice was made if the head of the lizard was inside the well of one of the ramps. We considered a trial failed if there was no choice in one hour of recording and those trials were considered as 'non data' in the analyses. We excluded from our analyses those individuals with more than 15 trials failed (i.e. they did not make a choice), and we considered the first trial to be the first one where the individual made a choice. We also considered individuals to be reinforced if the cricket was eaten after the test independently of whether their initial choice was correct or not.
